---
title: "2_CDL_DWR_USGS_crop_compare"
author: "Anna Boser"
date: "9/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(raster)
library(here)
library(tmap)
library(stringr)
```

## Comparison of land use data 

This script compares the land use data used in this study. These include: 
* the California Department of Water Resources 2018 crop map
* the Cropland Data Layer (this file uses 2018 for best comparison)
* the US Geographical Survey's county level irrigation data that also has acres of irrigated land 


### Read in data
```{r read_data}
# CDL (2018)
CDL <- raster(here("data", "raw", "CDL", "CDL2018", "2018_30m_cdls.img"))

# DWR
DWR <- read_sf(here("data", "raw", "DWR_crop", "i15_Crop_Mapping_2018.shp")) %>%
  st_transform(crs(CDL))

# USGS (2010 and 2015)
irrigation2015 <- read.csv(here("data", #this is data from 2015 (most recent)
                       "raw", 
                       "USGS_waterdata",
                       "2015",
                       "water_use"), sep = "\t")[-1,] #first row just tells you the size of the entries; remove

irrigation2010 <- read.csv(here("data", #this is data from 2010
                       "raw", 
                       "USGS_waterdata",
                       "2010",
                       "water_use"), sep = "\t")[-1,] #first row just tells you the size of the entries; remove

irrigation2015$year <- 2015
irrigation2010$year <- 2010

irrigation <- rbind(irrigation2015, irrigation2010)

rm(irrigation2010)
rm(irrigation2015)

# a shapefile of California counties; change to CDL crs
counties <- st_read(here("data", "raw", "shapefiles", "study_area_counties", "study_area_counties.shp")) %>%
  st_transform(crs(CDL))
```

```{r}
CIMIS_ETo <- read_sf(here("data", "raw", "shapefiles", "CIMIS_ETo", "Zones_eto_utm10_nad27.shp")) %>% 
  st_transform(crs(CDL))
```


### Clean data
```{r clean_data}
# crop CDL to California 
CDL <- crop(CDL, counties)

# turn USGS into a shapefile by merging with California counties
irrigation$NAME <- str_remove(irrigation$county_nm, " County") # remove the "County" suffix to each county name in order to match to county shapefile names
counties_irrigation <- base::merge(counties, irrigation, by = "NAME")

#for some reason the DWR geometry is in 3d with 0 for the z value so I use at_zm to get rid of it. 
DWR <- st_zm(DWR)
#turn the interesting classes into a factor
DWR$SUBCLASS2 <- as.factor(DWR$SUBCLASS2) 
DWR$CLASS2 <- as.factor(DWR$CLASS2)
```

### Plot the CDL and the DWR data
```{r}
# plot(CDL)
# plot(DWR[sample(nrow(DWR), 10000),"CLASS2"])

tm_shape(CDL) + 
  tm_raster() + 
  tm_shape(DWR[sample(nrow(DWR), 10000),"CLASS2"]) +
  tm_borders(col = "red") + 
  tm_shape(counties) + 
  tm_borders()
```


### Compare DWR and CDL
```{r}
# try for a small example
DWR_save <- DWR
DWR <- DWR[800:803,]
CDL_save <- CDL
CDL <- crop(CDL, DWR)
# plot(CDL)
# plot(DWR[,"CLASS2"])
tm_shape(CDL) + 
  tm_raster() + 
  tm_shape(DWR) +
  tm_borders(col = "red")



# make the polygon into the same raster 
DWR_raster <- rasterize(DWR, CDL, "CLASS2") 


# develop a crosswalk so that both rasters can be compared

# what is the percent accuracy of the CDL relative to the 

CDL <- CDL_save
DWR <- DWR_save
```

### If we resample CDL to ECOSTRESS (70m) resolution, get rid of non-pure pixels, and compare DWR again, is it better? 
```{r}

```

### What is the total area of ag land according to the CDL and DWR? How does that compare? 
```{r}

```

### How does DWR/CDL ag numbers compare to USGS total ag numbers? 
```{r}

```

