---
title: "counterfactual vs agriculture"
author: "Anna Boser"
date: '2022-03-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
library(here)
library(tidyr)
library(parallel)
```

```{r}
data <- fread(here("data", "for_analysis", "full_grid_not_tidy_cv.csv"), drop = c("PET0", "PET1", "PET5", "ET0", "ET1", "ET5"))
data$PET <- rowMeans(select(data, PET2, PET3, PET4))

#convert ET to mm
to_mm <- function(watts_m2, date){ #watts = J/s
  hours_of_daylight <- getSunlightTimes(date, 37.9833, -121.8677)$sunset - 
    getSunlightTimes(date, 37.9833, -121.8677)$sunrise # lat and lon are SJ valley coordinates from google maps. timezone not specified since we only care about the difference
  hours_of_daylight <- as.numeric(hours_of_daylight)
  kg = watts_m2*hours_of_daylight*3600*(1/2257)*(1/1000) # J/s * hours (h) * seconds in an hour (s/h) * latent heat of fusion (g/J) * conversion to kg AKA mm over a m2 (kg/g)
}

data$ET2 <- to_mm(data$ET2, as.Date("2019/06/01"))
data$ET3 <- to_mm(data$ET3, as.Date("2019/08/15"))
data$ET4 <- to_mm(data$ET4, as.Date("2019/10/15"))

data$ET <- rowMeans(select(data, ET2, ET3, ET4))

data <- data[agriculture == 1 | counterfactual == 1,]
data <- pivot_longer(data, cols = c("agriculture", "counterfactual"), names_to = "landcover")
# rename counterfactual
data <- filter(data, value == 1)
data$value = NULL
data <- filter(data, elevation >=-300, aspect >=-300, slope >=-300, ET >=0)
data$landcover <- ifelse(data$landcover == "counterfactual", "Natural Land", "Agriculture")

# if there is no aspect then the algorithm assigns it 1.570796400 or 4.712389 so we turn these into NA
data$aspect <- ifelse(data$aspect %in% c(1.570796400, 4.712389), NA, data$aspect)
data$slope <- data$slope*90 #in degrees
data$slope <- ifelse(data$slope>30, NA, data$slope) #in degrees

data$elevation <- data$elevation/1000
```

```{r}
names(data) <- c("x", "y", "Elevation", "Aspect", "Slope", "Soil quality", "PET2", "PET3", "PET4", "ET2", "ET3", "ET4", "PET", "ET", "Landcover")

data2 <- data %>% select(-PET2, -PET3, -PET4, -ET2, -ET3, -ET4) %>%
  pivot_longer(cols = c("Elevation", "Aspect", "Slope", "Soil quality", "PET", "ET"), names_to = "variable") 

data2$variable <- factor(data2$variable, levels = c("Elevation", "Aspect", "Slope", "Soil quality", "PET", "ET"))
```

```{r}
data2 %>%
  ggplot() + 
  geom_density(aes(x = value, fill = Landcover), lwd = 0, alpha = .4) + 
  scale_fill_manual(values=c("seagreen", "darkgoldenrod4")) +
  facet_wrap(vars(variable), scales = "free") + 
  theme_void()
```


```{r}
names(data) <- c("x", "y", "Elevation (km)", "Aspect (radians)", "Slope (degrees)", "Soil quality (CA storie index)", "PET2", "PET3", "PET4", "ET2", "ET3", "ET4", "PET (mm)", "ET (mm)", "Landcover")

data2 <- data %>% select(-PET2, -PET3, -PET4, -ET2, -ET3, -ET4) %>%
  pivot_longer(cols = c("Elevation (km)", "Aspect (radians)", "Slope (degrees)", "Soil quality (CA storie index)", "PET (mm)", "ET (mm)"), names_to = "variable") 

data2$variable <- factor(data2$variable, levels = c("Elevation (km)", "Aspect (radians)", "Slope (degrees)", "Soil quality (CA storie index)", "PET (mm)", "ET (mm)"))

# labs <- c("Elevation (m)", "Aspect (radians)", "Slope (degrees)", "Soil quality (CA storie index)", "PET (mm)", "ET (mm)")
# names(labs) <- c("Elevation", "Aspect", "Slope", "Soil quality", "PET", "ET")

data2 %>%
  ggplot() + 
  geom_density(aes(x = value, fill = Landcover), lwd = 0, alpha = .4) + 
  scale_fill_manual(values=c("seagreen", "darkgoldenrod4")) +
  facet_wrap(vars(variable), scales = "free", labeller = labeller(Landcover = labs)) + 
  theme_classic() +
  labs(xlab = "", 
       ylab = "")
  # theme(axis.title.x=element_blank(),
  #       axis.title.y=element_text("Density"), 
  #       axis.text.y=element_blank(), 
  #       axis.ticks.y=element_blank())
```


# how far is the furthest agriculture from counterfactual? 
```{r}
dist <- fread(here("data", "outputs", "distance_ag_counter.csv"))
```

```{r}
ggplot(dist) + geom_histogram(aes(x=mindist))
ggplot(dist) + geom_point(aes(x=x, y=y, col=mindist))
```

One degree lat/lon is about 100km. I know that that gives me real bad estimates. 10km would also be bad. 1km would definitely be good, and that's the average distance... 

```{r}
summary(dist)
```


Workflow:
1. Take a random sample of 1/1000 of ag points and find the distance to their closest counterfactual pixel
2. optimize hyperparameters for the order of magnitude of the median
3. test on groupings with the size of the order of magnitude of the maximum, median, and mean distance. 


