---
title: "Plot validation"
author: "Anna Boser"
date: '2022-03-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(data.table)
library(here)
library(dplyr)
library(sf)
library(tmap)
library(latex2exp)
library (rgdal)
library(maptools)
```

```{r}
# function that reads and binds all of a df type
load_bind <- function(string){
  overall_stats <- list.files(here("data", "for_analysis", "spatial_validation"), full = TRUE, pattern = string)
  read_add_dist_col <- function(file){
    df <- fread(file)
    df$dist <- as.numeric(str_extract(file, paste0("(?<=", string, "_)[0-9]*(?=.csv)")))
    return(df)
  }
  overall_stats <- rbindlist(lapply(overall_stats, read_add_dist_col))
  return (overall_stats)
}

general <- load_bind("sklearn_RF_cv_validation_stats")
monthly <- load_bind("sklearn_RF_cv_test_stats_by_month")
fold <- load_bind("sklearn_RF_cv_fold_stats")
```


# how far is the furthest agriculture from counterfactual? 

Workflow:
1. Take a random sample of 1/1000 of ag points and find the distance to their closest counterfactual pixel
2. optimize hyperparameters for the order of magnitude of the median
3. test on groupings with the size of the order of magnitude of the maximum, median, and mean distance. 

```{r}
dist <- fread(here("data", "outputs", "distance_ag_counter.csv"))
```

```{r}
coef = 2/5

ggplot(NULL) + 
  stat_density(data = dist, aes(x=mindist/1000), fill = "blue") + 
  geom_line(data = general, aes(x=dist/1000, y = r2*coef), color = "red") +
  geom_point(data = general, aes(x=dist/1000, y = r2*coef), color = "red") +
  xlab(TeX("Distance of agricultural pixel to nearest natural pixel (km) \n Size of areas left out of model training (km^2)")) + 
    theme_classic() + 
  theme(axis.title.x = element_text(vjust=-2.5)) +
  scale_y_continuous(
    name = "Distribution of agricultural pixels",
    sec.axis = sec_axis(~./coef, name=TeX("Model performance ($R^{2}$)")))

```

what about different months? 
```{r}
ggplot(monthly) + 
  geom_line(aes(x=dist/1000, y = r2, color = monthgroup, group = monthgroup)) +
  geom_point(aes(x=dist/1000, y = r2, color = monthgroup))
```

what about plotting over space? 
```{r}
# first turn "fold" into x and y coordinates
fold$x <- as.numeric(str_extract(fold$cv_fold, "[-.0-9]*(?=,)"))
fold$y <- as.numeric(str_extract(fold$cv_fold, "(?<=,)[-.0-9]*"))

# get the central valley
CV <- readOGR(here("data", "raw", "shapefiles", "cimis_CV", "cimis_CV.shp"))
CV <- spTransform(CV, CRS("+proj=longlat +datum=WGS84"))
CV <- fortify(CV)
```

```{r}
filter(fold) %>%
  ggplot() + 
  geom_raster(aes(x=x, y=y, fill=r2)) +
  facet_wrap(vars(dist/1000)) +
  # scale_fill_gradientn(colours = c("darkgoldenrod4", "darkgoldenrod3", "darkgoldenrod2", "darkgoldenrod1", "khaki1", "lightgreen", "turquoise3", "deepskyblue3", "mediumblue", "mediumblue", "navyblue", "navyblue", "midnightblue", "midnightblue", "midnightblue", "black", "black")) +
  # scale_fill_gradientn(colours = c("darkgoldenrod4", "darkgoldenrod2", "khaki1", "lightgreen", "turquoise3", "deepskyblue3", "mediumblue", "navyblue", "midnightblue", "black")) +
  scale_fill_distiller(palette="Spectral", direction = 1) +
  geom_polygon(data = CV, aes(long, lat, group=group), colour = alpha("black", 1), fill = NA) +
  theme_classic() 
```

































```{r}
data$lon <- floor(data$cv_fold/1000)
data$lat <- data$cv_fold - data$lon*1000
```

```{r}
CV_grid <- raster(here("data", "intermediate", "CA_grid_cv.tif"))
CA <- st_read(here("data", "raw", "shapefiles", "california", "california.shp")) %>% st_transform(st_crs(CV_grid))
CV <- st_read(here("data", "raw", "shapefiles", "cimis_CV", "cimis_CV.shp")) %>% st_transform(st_crs(CV_grid))

# make a raster of 1x1 degree cells
raster = raster(#nrows=10, 
                #ncols=11, 
                xmn=min(data$lon), 
                xmx=max(data$lon) + 1, # because of the flooring operation
                ymn=min(data$lat), 
                ymx=max(data$lat) + 1, 
                crs = st_crs(CV), 
                resolution = 1, 
                vals=1:35)

# turn it into a shapefile since that'll be easier
sf = rasterToPolygons(raster)
sf <- st_as_sf(sf)
st_crs(sf) <- st_crs(CV)

```

```{r}
# crs(sf) <- st_crs(CV)
tm_shape(sf) + tm_polygons(col="layer") + tm_shape(CV) + tm_borders()
```

```{r}
# remove the ones that do not overlap with the central valley
sf1 <- sf[CV,]
tm_shape(sf1) + tm_polygons(col="layer") + tm_shape(CV) + tm_borders()
```


```{r}
head(data)
```


```{r}
data %>% 
  ggplot() + 
  geom_point(aes(x = ET, y= ET_pred, col = lat), alpha = .2, size = .1)
```

```{r}
data %>% 
  ggplot() + 
  geom_point(aes(x = ET, y= ET_pred, col = factor(monthgroup)), alpha = .2, size = .1)
```


```{r}
data %>% 
  group_by(lon, lat) %>%
  summarise(rmse = mean((ET_pred-ET)^2), 
            mae = mean(abs(ET_pred-ET))) %>%
  ggplot() +
  geom_point(aes(x = lon, y = lat, col = mae))
  
  
```

