---
title: "irrigation comparison"
author: "Anna Boser"
date: '2022-05-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(stringr)
library(sf)
library(tmap)
library(data.table)
library(tidyr)
library(dplyr)
library(ggplot2)
library(rgdal)
library(ggnewscale) # allows you to plot multiple things with different scales on one ggplot
```


## ET data

```{r}
# read ET data
data <- fread(file = here("data/for_analysis/ag_counterfactual/ag_counterfactual_default0.1.csv"))
```

```{r}
# read county information and merge with ET data
counties <- fread(file=file.path(here("data", "intermediate", "counties", "counties.csv")))
counties$x <- round(counties$x, 7)
counties$y <- round(counties$y, 7)
data$x <- round(data$x, 7)
data$y <- round(data$y, 7)
data <- merge(data, counties, by = c("x", "y"))
rm(counties)
```

I need to average over monthgroups
```{r}
data <- pivot_wider(data, names_from = c(monthgroup), values_from = c(ET, ET_pred, ag_ET, PET))

data$ET <- rowMeans(select(data, ET_2, ET_3, ET_4))
data$ET_pred <- rowMeans(select(data, ET_pred_2, ET_pred_3, ET_pred_4))
data$ag_ET <- rowMeans(select(data, ag_ET_2, ag_ET_3, ag_ET_4))
data$PET <- rowMeans(select(data, PET_2, PET_3, PET_4))
```

I then need to get the total water ETd, so I average over county and then multiply by the number of days in my observation window: 183
```{r}
county_data <- data %>% 
  group_by(NAME) %>% # group by county here
  summarize(
    ET_sd = sd(ET, na.rm=TRUE), 
    ET = mean(ET, na.rm=TRUE), 
    ET_pred_sd = sd(ET_pred, na.rm=TRUE),
    ET_pred = mean(ET_pred, na.rm=TRUE), 
    ag_ET_sd = sd(ag_ET, na.rm=TRUE), 
    ag_ET = mean(ag_ET, na.rm=TRUE), 
    PET_sd = sd(PET, na.rm=TRUE), 
    PET = mean(PET, na.rm=TRUE), 
    n = n()) #get the number of pixels in each county

county_data$ET_year <- county_data$ET*183
county_data$ET_pred_year <- county_data$ET_pred*183
county_data$PET_year <- county_data$PET*183
county_data$ag_ET_year <- county_data$ag_ET*183
```


## USGS data

### Read in USGS data. Make a single irrigation dataset with 2010 and 2015 data
```{r}
irrigation2015 <- read.csv(here("data", #this is data from 2015 (most recent)
                       "raw", 
                       "USGS_waterdata",
                       "2015",
                       "water_use"), sep = "\t")[-1,] #first row just tells you the size of the entries; remove

irrigation2010 <- read.csv(here("data", #this is data from 2010
                       "raw", 
                       "USGS_waterdata",
                       "2010",
                       "water_use"), sep = "\t")[-1,] #first row just tells you the size of the entries; remove

irrigation2015$year <- 2015
irrigation2010$year <- 2010

irrigation <- rbind(irrigation2015, irrigation2010)

rm(irrigation2010)
rm(irrigation2015)
```

### Clean USGS irrigation data
Change units to kg/day and meters squared
```{r}
irrigation$NAME <- str_remove(irrigation$county_nm, " County") # remove the "County" suffix to each county name in order to match to county shapefile names


# convert billions of gallons per day to kg per day
conversion <- function(Mgal.d){
  Mgal.d = as.numeric(Mgal.d)
  kg.d = Mgal.d * 3.785411784 * 1000000
}

to_convert = names(irrigation)[c(6:10, 15)] #the names of the columns that are in Mgal/d and need to be converted
irrigation <- mutate(irrigation, across(to_convert, conversion))


# convert thousands of acres to square meters
conversion <- function(tacre){
  tacre = as.numeric(tacre)
  m2 = tacre * 1000 * 4046.86 # conversion to acres and conversion to square meters
}

to_convert = names(irrigation)[c(11:14)] #the names of the columns that are in thousand of acres and need to be converted

irrigation <- mutate(irrigation, across(to_convert, conversion))

# add a mm/day and mm/year variable
# keep in mind the column names no longer reflect the true units
irrigation$mm.day <- irrigation$Irrigation..Crop.total.self.supplied.withdrawals.for.crops..fresh..in.Mgal.d/irrigation$Irrigation..Crop.total.irrigation.for.crops..in.thousand.acres

irrigation$mm.year <- irrigation$mm.day*365
```

### Calculate the irrigation efficiency
```{r}
# average across 2010 and 2015 irrigation data
irrigation_avg <- irrigation %>% 
  group_by(NAME) %>% 
  summarize(mm.year = mean(mm.year))

#get a yearly estimate of efficiency
data$ag_ET_year <- data$ag_ET*183

# list of counties to include -- basically the ones that are sufficiently in the CV
cv_counties <- unique(data$NAME) # the ones with ag in the CV
cv_counties <- cv_counties[!(cv_counties %in% c("Shasta", 
                                                "Nevada", 
                                                "El Dorado", 
                                                "Amador", 
                                                "Solano", 
                                                "Alameda", 
                                                "Contra Costa", 
                                                "Calaveras", 
                                                "Tuolomne", 
                                                "Mariposa"))] # remove a few that don't have enough ag in the CV

cv_irrigation <- filter(irrigation_avg, NAME %in% cv_counties)
cv_data <- filter(data, NAME %in% cv_counties)

# merge the two, calculate efficiency for each plot, and then take the average
cv_efficiency <- base::merge(cv_irrigation, cv_data, by = "NAME", all.y = TRUE, all.x = FALSE)
cv_efficiency$efficiency <- cv_efficiency$ag_ET_year/cv_efficiency$mm.year
efficiency <- mean(cv_efficiency$efficiency, na.rm = TRUE) # 20%!
```


### Read in county shapefile
```{r}
counties <- st_read(here("data", "raw", "shapefiles", "study_area_counties", "study_area_counties.shp"))
```

### Merge dataframes 
```{r}
counties_irrigation <- base::merge(counties, irrigation, by = "NAME")
```

### Plot irrigation data -- mm over agricultural lands for 2010 and 2015
```{r}
# 2010
filter(counties_irrigation, year == 2010) %>% 
  tm_shape() +
  tm_polygons("mm.year")

# 2015
filter(counties_irrigation, year == 2015) %>% 
  tm_shape() +
  tm_polygons("mm.year")
```

## Both datasets

```{r}
CV <- st_read(here("data", "raw", "shapefiles", "cimis_CV", "cimis_CV.shp"))
```


```{r}
# take the average irrigation over different years to simplify
counties_irrigation_avg <- counties_irrigation %>% 
  group_by(NAME) %>% 
  summarize(mm.year = mean(mm.year))

county_data <- base::merge(counties_irrigation_avg, county_data, by="NAME")
county_data$irrigation_efficiency <- county_data$ag_ET_year*100/county_data$mm.year
```


```{r}
county_data %>% 
  tm_shape() +
  tm_polygons("irrigation_efficiency") + 
  tm_shape(CV) + 
  tm_borders()
```


## try plotting with ggplot -- see https://medium.com/@honggao/plot-polygons-with-ggplot2-b5a87e4fa602
```{r}
ggplot() + 
  geom_sf(data = county_data, aes(fill = irrigation_efficiency), color=alpha("red",0)) + 
  scale_fill_gradient2(low = "red", mid = "white", high = "#70AD47") + 
  geom_sf(data = CV, fill=alpha("red",0)) + 
  new_scale_fill() +
  geom_raster(data = data, aes(x=x, y=y, fill=ag_ET), alpha=0.75) +
  scale_fill_gradientn(colours = c("darkgoldenrod4", "darkgoldenrod2", "khaki1", "lightgreen", "turquoise3", "deepskyblue3", "mediumblue", "navyblue", "midnightblue", "black"), limits = c(-2, 8)) +
  geom_sf(data = county_data, fill=alpha("red",0), color=alpha("grey64")) + 
  theme_void()
```

```{r}
ggplot() + 
  geom_sf(data = county_data, aes(fill = irrigation_efficiency), color=alpha("red",0)) + 
  scale_fill_gradient2(low = "red", mid = "white", high = "#70AD47") + 
  geom_sf(data = CV, fill=alpha("red",0)) + 
  theme_void()
```

<!-- ```{r} -->
<!-- data %>%  -->
<!--   tm_shape() + -->
<!--   tm_polygons("ag_ET") +  -->
<!--   tm_shape(CV) +  -->
<!--   tm_borders() -->

<!-- data %>%  -->
<!--   tm_shape() + -->
<!--   tm_polygons("ag_ET_sd") +  -->
<!--   tm_shape(CV) +  -->
<!--   tm_borders() -->

<!-- data %>%  -->
<!--   tm_shape() + -->
<!--   tm_polygons("n") +  -->
<!--   tm_shape(CV) +  -->
<!--   tm_borders() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- filter(data, year == 2010) %>%  -->
<!--   tm_shape() + -->
<!--   tm_polygons("mm.year") +  -->
<!--   tm_shape(CV) +  -->
<!--   tm_borders() -->

<!-- filter(data, year == 2015) %>%  -->
<!--   tm_shape() + -->
<!--   tm_polygons("mm.year") +  -->
<!--   tm_shape(CV) +  -->
<!--   tm_borders() -->
<!-- ``` -->


<!-- ```{r} -->
<!-- data$percent_total_ET <- data$ET_year*100/data$mm.year -->

<!-- filter(data, year == 2010) %>%  -->
<!--   tm_shape() + -->
<!--   tm_polygons("percent_total_ET") +  -->
<!--   tm_shape(CV) +  -->
<!--   tm_borders() -->

<!-- filter(data, year == 2015) %>%  -->
<!--   tm_shape() + -->
<!--   tm_polygons("percent_total_ET") +  -->
<!--   tm_shape(CV) +  -->
<!--   tm_borders() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- data %>%  -->
<!--   tm_shape() + -->
<!--   tm_polygons("ET_year") -->
<!-- ``` -->


<!-- Irrigation efficiency and irrigation technology -->
<!-- Let's do some analysis of types of irrigation versus efficiency when controlling for PET -->
<!-- ```{r} -->
<!-- data <- data %>%  -->
<!--   mutate( -->
<!--     sprinkler = Irrigation..Crop.sprinkler.irrigation.for.crops..in.thousand.acres/Irrigation..Crop.total.irrigation.for.crops..in.thousand.acres,  -->
<!--     microirrigation = Irrigation..Crop.microirrigation.for.crops..in.thousand.acres/Irrigation..Crop.total.irrigation.for.crops..in.thousand.acres,  -->
<!--     surface = Irrigation..Crop.surface.irrigation.for.crops..in.thousand.acres/Irrigation..Crop.total.irrigation.for.crops..in.thousand.acres) -->

<!-- # data2015 %>% pivot_longer(cols = c(sprinkler, microirrigation, surface), names_to="irrigation_type", values_to = "percent") -->
<!-- ``` -->


<!-- ```{r} -->
<!-- lm <- lm(irrigation_efficiency~sprinkler+microirrigation+surface -1, data) -->
<!-- summary(lm) -->
<!-- ``` -->


<!-- microirrigation and PET are inversely correlated, meaning micro is probably being undervalued, while surface is positively correlated, meaning it's likely overvalued.  -->
<!-- ```{r} -->
<!-- lm <- lm(irrigation_efficiency~sprinkler+microirrigation+surface+PET -1, data) -->
<!-- summary(lm) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- data %>%  -->
<!--   pivot_longer(cols = c(sprinkler, microirrigation, surface), names_to="irrigation_type", values_to = "percent") %>% -->
<!--   ggplot(aes(x=percent, y=irrigation_efficiency)) +  -->
<!--   geom_point(aes(color=NAME)) +  -->
<!--   geom_smooth(method="lm") + -->
<!--   facet_grid(cols = vars(irrigation_type)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- data %>%  -->
<!--   ggplot(aes(x=PET, y=irrigation_efficiency)) +  -->
<!--   geom_point() +  -->
<!--   geom_smooth(method="lm") -->
<!-- ``` -->
<!-- ^statistically significant -->

<!-- ```{r} -->
<!-- lm <- lm(ag_ET~sprinkler+microirrigation+surface -1, filter(data, year == 2015)) -->
<!-- summary(lm) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- lm <- lm(ag_ET~surface+microirrigation+PET, filter(data, year == 2015)) -->
<!-- summary(lm) -->
<!-- ``` -->
